Emulation of high multiplicity matrix elements for electron-positron and hadron-hadron colliders

### 2022-09-23 appendix1.tex
wrote: matrix element factorisation
wrote: colour-ordered and one-loop factorisation

\documentclass[main.tex]{subfiles}
\begin{document}
\chapter{Conclusion}
\label{chapter:conclusion}

>>>>

\documentclass[main.tex]{subfiles}
\begin{document}
\chapter{Catani-Seymour notation}
\label{appendix:catani_seymour}
    The notation for matrix elements given in Ref.~\cite{Catani:1996vz}
    is repeated here for convenience.

    Consider a tree-level matrix element with $n$ partons in
    the final-state
    \begin{equation}\label{eqn:CS_ME}
        \mathcal{M}^{c_{1},\ldots,c_{n};s_{1},\ldots,s_{n}}_{n}(p_{1},\ldots,p_{n}) \, ,
    \end{equation}
    where $\{c_{1},\ldots,c_{n}\}$ are colour indices,
    $\{s_{1},\ldots,s_{n}\}$ are spin indices,
    and $\{p_{1},\ldots,p_{n}\}$ are momenta.

    Introducing the basis $\{| c_{1},\ldots,c_{n} \rangle \otimes | s_{1},\ldots,s_{n} \rangle \}$
    in colour + helicity space allows us to write
    \begin{equation}\label{eqn:CS_hilbert}
        \mathcal{M}^{c_{1},\ldots,c_{n};s_{1},\ldots,s_{n}}_{n}(p_{1},\ldots,p_{n}) \equiv \left(\langle c_{1},\ldots,c_{n} | \otimes \langle s_{1},\ldots,s_{n} | \right) | 1, \ldots,n \rangle_{n} \, ,
    \end{equation}
    where $|1, \ldots, n \rangle_{n}$ is a vector in colour + helicity space.
    The indices inside the vector denote the spin and colour indices,
    while the subscript denotes the multiplicity of the vector.
    The matrix element squared, summed over final-state colours and spins,
    can therefore be written as
    \begin{equation}\label{eqn:CS_ME2}
        |\mathcal{M}_{n}|^{2} = {}_{n} \langle 1,\ldots,n|1,\ldots,n\rangle_{n} \, .
    \end{equation}

    In the case of partons in the initial-state the colour + helicity vector
    $|1, \ldots, n \rangle$ has to be normalised by a factor $\sqrt{n_{c}}$ for each
    initial state parton. For hadron-hadron collisions this amounts to
    \begin{equation}
        | 1, \ldots, n \rangle_{n} \rightarrow \dfrac{1}{\sqrt{n_{c}(a)n_{c}(b)}} | 1, \ldots, n \rangle_{n} \, ,
    \end{equation}
    where $n_{c}(a)$ denotes the number of colour states of parton $a$.
    Namely $n_{c}(q) = n_{c}(\bar{q}) = N_{c}$ and $n_{c}(g) = N_{c}^{2}-1$.

### 2022-09-23 qcd.tex
wrote: colour-ordered and one-loop factorisation

    The behaviour of $\alpha_{s}$
    is governed by the Callan-Symanzik \cite{Callan:1970yg,Symanzik:1970rt}
    $\beta$-function
    \begin{equation}\label{eqn:beta_fn}
        \mu^{2} \dfrac{\partial \alpha_{s}(\mu^{2})}{\partial \mu^{2}} = \beta(\alpha_{s}) \, ,
    \end{equation}
    where the $\beta$-function can be written as a perturbative
    expansion in $\alpha_{s}$
    \begin{equation}\label{eqn:beta_series}
        -\beta(\alpha_{s}) = \alpha_{s}\sum_{n=0}^{\infty}\left(\dfrac{\alpha_{s}}{4\pi}\right)^{n+1} \beta_{n}\, 
    \end{equation}
    where the coefficients $\beta_{n}$ have been computed
    up to $\beta_{4}$ \cite{Baikov:2016tgj,Luthe:2017ttg}.
    The solution to (\ref{eqn:beta_fn}) to first order
    is
    \begin{equation}\label{eqn:1l_alpha}
        \alpha_{s}(\mu^{2}) = \dfrac{1}{\frac{\beta_{0}}{4\pi}\log\left(\frac{\mu^{2}}{\Lambda_{\mathrm{QCD}}^{2}}\right)} \, ,
    \end{equation}
    where $\Lambda_{\mathrm{QCD}}$ is the QCD scale,
    the scale at which $\alpha_{s}$ becomes
    large enough that perturbation theory breaks down,
    and $\beta_{0} = (11C_{A} - 4T_{R}n_{f})/3$.
    In QCD, $C_{A}=3$ and $T_{R}=\frac{1}{2}$, meaning for
    $n_{f} < \frac{33}{2}$ the sign of $\beta_{0}$ is positive,
    which is what we have observed in nature with six
    flavours of quarks. This behaviour of positive
    $\beta_{n}$ is also the case in all coefficients
    computed to date, meaning that $\alpha_{s}$
    decreases with increasing energy due to the minus sign
    in (\ref{eqn:beta_series}). This property is known as
    asymptotic freedom \cite{Gross:1973id,Politzer:1973fx},
    and is justification for treating QCD perturbatively.
    The running of the coupling constant has been
    observed experimentally as illustrated in Figure~\ref{fig:alpha_s_running}.
    \begin{figure}
        \includegraphics{qcd/alpha_s_running.pdf}
        \caption{The running of the strong coupling constant $\alpha_{s}$,
        as determined by experiments, with QCD theory prediction in black.
        Figure reference \cite{Workman:2022ynf}.}
        \label{fig:alpha_s_running}
    \end{figure}

@@

    Schematically, the $(n+1)$-body matrix element
    factorises as
    \begin{equation}\label{eqn:me_factorisation}
        |\mathcal{M}_{n+1}|^{2} \rightarrow S_{ijk} \otimes |\mathcal{M}_{n}|^{2} \, ,
    \end{equation}
    where $S_{ijk}$ is a universal singular factor capturing
    the IR divergent behaviour and $|\mathcal{M}_{n}|^{2}$
    is the reduced matrix element. The $\otimes$
    represents the colour and spin correlations existing
    between the singular function and the reduced matrix element.
    This singular factor
    is not unique and can be represented by different
    approximations as long as they reproduce the
    correct IR behaviour. In this thesis we discuss
    two approximations: Catani-Seymour dipoles \cite{Catani:1996vz}
    in Section~\ref{sec:CS_dipoles} and antenna functions \cite{Gehrmann-DeRidder:2005btv}
    in Section~\ref{sec:antenna_functions}. The indices
    of $S_{ijk}$ already hints at the fact that the singular
    functions only depends on three partons, and not all
    of the $n+1$ external states.

    We will inspect the soft and collinear limits separately
    to see how the matrix element factorises in these respective
    limits.
    The following section will follow the conventions
    of Ref.~\cite{Catani:1996vz}, namely that the matrix
    element for $n$ partons in the final-state is written as
    \begin{equation}\label{eqn:CS_matrix_element}
        |\mathcal{M}_{n}|^{2} = {}_{n}\langle 1, \ldots, n | 1, \ldots, n \rangle_{n}
    \end{equation}
    where $| 1, \ldots, n \rangle_{n}$ is a vector
    in colour + helicity space (see Appendix~\ref{appendix:catani_seymour}
    for a more thorough explanation on notation).

\subsection{Soft limits}\label{sec:me_soft}
    Consider a tree-level matrix element $|\mathcal{M}_{n+1}|^{2}$
    with a final-state gluon $j$. Note that reduced matrix
    elements associated with taking a quark soft have to vanish
    due to violation of quark number, therefore only gluons
    are considered.
    The limit of the soft gluon with momentum
    $p_{j}$ can be parametrised by
    \begin{equation}\label{eqn:soft_parametrisation}
        p_{j}^{\mu} = \lambda q^{\mu} \, , \quad \lambda \rightarrow 0 \, ,
    \end{equation}
    where $q^{\mu}$ is an arbitrary four-vector and $\lambda$
    is a scale parameter. In this limit the matrix element
    can be written as
    \begin{equation}\label{eqn:soft_factorisation}
        |\mathcal{M}_{n+1}|^{2} \rightarrow -\dfrac{1}{\lambda^{2}}8\pi \mu^{2\epsilon}\alpha_{s} \sum_{i}\dfrac{1}{p_{i}q}\sum_{k \neq i} \dfrac{p_{k}p_{i}}{(p_{i} + p_{k})q} {}_{n}\langle 1, \ldots, n+1 | \bm{T}_{k} \cdot \bm{T}_{i} | 1, \ldots, n+1 \rangle_{n} \, ,
    \end{equation}
    where terms less singular than $1 / \lambda^{2}$ have been
    neglected. The indices $i$, $j$, and $k$ label partons involved
    in the factorsation process: $i$ is the emitter parton,
    $j$ is the emitted parton, and $k$ is a parton accounting
    for the colour correlations. The scale $\mu$ can be identified as the
    renormalisation scale, and $\bm{T_{i}}$ is the colour-charge
    operator. It acts on the colour vectors to give factors of
    $C_{A}$ and $C_{F}$, hence this factorisation is not exact.

@@

\subsection{Collinear limits}\label{sec:me_collinear}
    For the collinear limit, consider partons $i$ and
    $j$ in the matrix element $|\mathcal{M}_{n+1}|^{2}$.
    Their momenta can be decomposed as
    \begin{align}\label{eqn:collinear_momenta}
        \begin{split}
        p_{i}^{\mu} &= z p^{\mu} + k_{\perp}^{\mu} - \dfrac{k_{\perp}^{2}}{z} \dfrac{n^{\mu}}{2 p n} \, \\
        p_{j}^{\mu} &= (1-z) p^{\mu} - k_{\perp}^{\mu} - \dfrac{k_{\perp}^{2}}{1-z}\dfrac{n^{\mu}}{2 p n} \, , 
        \end{split}
    \end{align}
    where $p^{\mu}$ denotes the collinear direction
    of the two partons, $k_{\perp}^{\mu}$ specifies
    the transverse direction perpendicular to the
    collinear direciton, and $n^{\mu}$ is an auxiliary
    vector satisfying the conditions $n^{2} = 0$ and $k_{\perp} n = 0$.
    $z$ is the fraction of momenta carried away from
    the collinear momentum by parton $i$.

    The collinear limit can be then be defined as
    \begin{equation}\label{eqn:collinear limit}
        2 p_{i} p_{j} = - \dfrac{k_{\perp}^{2}}{z(1-z)} \, , \quad k_{\perp} \rightarrow 0 \, .
    \end{equation}
\section{Subtraction}\label{sec:subtraction}
\subsection{Catani-Seymour dipoles}\label{sec:CS_dipoles}
\subsection{Antenna}\label{sec:antenna_functions}
\end{document}

>>>>

    The behaviour of $\alpha_{s}$
    is governed by the Callan-Symanzik \cite{Callan:1970yg,Symanzik:1970rt}
    $\beta$-function
    \begin{equation}\label{eqn:beta_fn}
        \mu^{2} \dfrac{\partial \alpha_{s}(\mu^{2})}{\partial \mu^{2}} = \beta(\alpha_{s}) \, ,
    \end{equation}
    where the $\beta$-function can be written as a perturbative
    expansion in $\alpha_{s}$
    \begin{equation}\label{eqn:beta_series}
        -\beta(\alpha_{s}) = \alpha_{s}\sum_{n=0}^{\infty}\left(\dfrac{\alpha_{s}}{4\pi}\right)^{n+1} \beta_{n}\, 
    \end{equation}
    where the coefficients $\beta_{n}$ have been computed
    up to $\beta_{4}$ \cite{Baikov:2016tgj,Luthe:2017ttg}.
    The solution to (\ref{eqn:beta_fn}) to first order
    is
    \begin{equation}\label{eqn:1l_alpha}
        \alpha_{s}(\mu^{2}) = \dfrac{1}{\frac{\beta_{0}}{4\pi}\log\left(\frac{\mu^{2}}{\Lambda_{\mathrm{QCD}}^{2}}\right)} \, ,
    \end{equation}
    where $\Lambda_{\mathrm{QCD}}$ is the QCD scale,
    the scale at which $\alpha_{s}$ becomes
    large enough that perturbation theory breaks down,
    and $\beta_{0} = (11C_{A} - 4T_{R}n_{f})/3$.
    In QCD, $C_{A}=3$ and $T_{R}=\frac{1}{2}$, meaning for
    $n_{f} < \frac{33}{2}$ the sign of $\beta_{0}$ is positive,
    which is what we have observed in nature with six
    flavours of quarks. This behaviour of positive
    $\beta_{n}$ is also the case in all coefficients
    computed to date, meaning that $\alpha_{s}$
    decreases with increasing energy due to the minus sign
    in (\ref{eqn:beta_series}). This property is known as
    asymptotic freedom \cite{Gross:1973id,Politzer:1973fx},
    and is justification for treating QCD perturbatively.
    The running of the coupling constant has been
    observed experimentally as illustrated in Figure~\ref{fig:alpha_s_running}.

    \begin{figure}
        \includegraphics{qcd/alpha_s_running.pdf}
        \caption{The running of the strong coupling constant $\alpha_{s}$,
        as determined by experiments, with QCD theory prediction in black.
        Figure reference \cite{Workman:2022ynf}.}
        \label{fig:alpha_s_running}
    \end{figure}

@@

    Schematically, the $(n+1)$-body matrix element
    factorises as
    \begin{equation}\label{eqn:me_factorisation}
        |\mathcal{M}_{n+1}|^{2} \rightarrow \bm{S}_{ijk} \otimes |\mathcal{M}_{n}|^{2} \, ,
    \end{equation}
    where $\bm{S}_{ijk}$ is a universal singular factor capturing
    the IR divergent behaviour and $|\mathcal{M}_{n}|^{2}$
    is the reduced matrix element. The $\otimes$
    represents the colour and spin correlations existing
    between the singular function and the reduced matrix element.
    This singular factor
    is not unique and can be represented by different
    approximations as long as they reproduce the
    correct IR behaviour. In this thesis we discuss
    two approximations: Catani-Seymour dipoles \cite{Catani:1996vz}
    in Section~\ref{sec:CS_dipoles} and antenna functions \cite{Gehrmann-DeRidder:2005btv}
    in Section~\ref{sec:antenna_functions}. The indices
    of $\bm{S}_{ijk}$ already hints at the fact that the singular
    functions only depends on three partons, and not all
    of the $n+1$ external states.

    We will inspect the soft and collinear limits separately
    to see how the matrix element factorises in these respective
    limits.
    The following section will follow the conventions
    of Ref.~\cite{Catani:1996vz}, namely that the colour and
    helicity summed $n$-body matrix element written as
    \begin{equation}\label{eqn:CS_matrix_element}
        |\mathcal{M}_{n}|^{2} = {}_{n}\langle 1, \ldots, n | 1, \ldots, n \rangle_{n}
    \end{equation}
    where $| 1, \ldots, n \rangle_{n}$ is a vector
    in colour + helicity space (see Appendix~\ref{appendix:catani_seymour}
    for a more thorough explanation on notation).

\subsection{Soft limits}\label{sec:me_soft}
    Consider a tree-level matrix element $|\mathcal{M}_{n+1}|^{2}$
    with a final-state gluon $j$. Note that reduced matrix
    elements associated with taking a quark soft have to vanish
    due to violation of quark number, therefore only gluons
    are considered.
    The limit of the soft gluon with momentum
    $p_{j}$ can be parametrised by
    \begin{equation}\label{eqn:soft_parametrisation}
        p_{j}^{\mu} = \lambda q^{\mu} \, , \quad \lambda \rightarrow 0 \, ,
    \end{equation}
    where $q^{\mu}$ is an arbitrary four-vector and $\lambda$
    is a scale parameter. In this limit the matrix element
    can be written as
    \begin{equation}\label{eqn:soft_factorisation}
        |\mathcal{M}_{n+1}|^{2} \rightarrow -\dfrac{1}{\lambda^{2}}8\pi \mu^{2\epsilon}\alpha_{s} \sum_{i}\dfrac{1}{p_{i}q}\sum_{k \neq i} \dfrac{p_{k}p_{i}}{(p_{i} + p_{k})q} \; {}_{n}\langle 1, \ldots, n+1 | \bm{T}_{k} \cdot \bm{T}_{i} | 1, \ldots, n+1 \rangle_{n} \, ,
    \end{equation}
    where terms less singular than $1 / \lambda^{2}$ have been
    neglected. The reduced matrix element is obtained by removing
    the soft gluon. The indices $i$, $j$, and $k$ label partons involved
    in the factorsation process: $i$ is the emitter parton,
    $j$ is the emitted parton, and $k$ is a parton accounting
    for the colour correlations. The scale $\mu$ can be identified as the
    renormalisation scale, and $\bm{T_{i}}$ is the colour-charge
    operator. It acts on the colour vectors to give factors of
    $C_{A}$ and $C_{F}$, hence this factorisation is not exact.

@@

\subsection{Collinear limits}\label{sec:me_collinear}
    For the collinear limit, consider partons $i$ and
    $j$ in the matrix element $|\mathcal{M}_{n+1}|^{2}$.
    Their momenta can be decomposed as
    \begin{align}\label{eqn:F_collinear_momenta}
        \begin{split}
        p_{i}^{\mu} &= z p^{\mu} + k_{\perp}^{\mu} - \dfrac{k_{\perp}^{2}}{z} \dfrac{n^{\mu}}{2 p n} \, \\
        p_{j}^{\mu} &= (1-z) p^{\mu} - k_{\perp}^{\mu} - \dfrac{k_{\perp}^{2}}{1-z}\dfrac{n^{\mu}}{2 p n} \, , 
        \end{split}
    \end{align}
    where $p^{\mu}$ denotes the collinear direction
    of the two partons, $k_{\perp}^{\mu}$ specifies
    the transverse direction perpendicular to the
    collinear direciton ($p k_{\perp} = 0$), and $n^{\mu}$ is an auxiliary
    vector satisfying the conditions $n^{2} = 0$ and $k_{\perp} n = 0$.
    $z$ is the fraction of momenta carried away from
    the collinear momentum by parton $i$.

    The collinear limit can be then be defined as
    \begin{equation}\label{eqn:F_collinear limit}
        2 p_{i} p_{j} = s_{ij} = - \dfrac{k_{\perp}^{2}}{z(1-z)} \, , \quad k_{\perp} \rightarrow 0 \, .
    \end{equation}
    In this limit the matrix element can be written as
    \begin{equation}\label{eqn:F_collinear_factorisation}
        |\mathcal{M}_{n+1}|^{2} \rightarrow \dfrac{1}{p_{i}p_{j}}4\pi \mu^{2\epsilon}\alpha_{s} \; {}_{n} \langle 1, \ldots, n+1 | \hat{P}_{ij}(z, k_{\perp}) | 1, \ldots, n+1 \rangle_{n} \, ,
    \end{equation}
    where terms less singular than $k_{\perp}^{2}$ have been neglected.
    This reduced matrix element is obtained by replacing the partons
    $i$ and $j$ with a single parton $ij$ which carries the
    momentum $p^{\mu}$ and suitable quantum numbers depending
    on the partons $i$ and $j$. For example, if $i =$ quark and $j=$ gluon,
    then $ij =$ quark, or if $i=$ quark and $j =$ antiquark, then $ij =$ gluon.
    $\hat{P}_{ij}$ are the $d$-dimensional Altarelli-Parisi splitting
    functions that depend on the momentum fraction $z$ for the splitting
    $ij \rightarrow i + j$ and the transverse momentum $k_{\perp}$.
    Each splitting function is matrix acting on the spin indices of $ij$.
    Due to these spin correlations the matrix
    element does not factorise from the splitting functions exactly.

    $\hat{P}_{ij}$ become the more recognisable Altarelli-Parisi
    splitting functions (\ref{eqn:AP_kernels}) once spin-averaged
    and the $\epsilon \rightarrow 0$ limit is taken.

    In general, it is possible for the case in which a final-state
    parton $i$ becomes collinear with an initial-state parton $a$.
    This is described by the splitting process $a \rightarrow ai + i$.
    In this case their momenta can be decomposed as
    \begin{align}\label{eqn:I_collinear_momenta}
        \begin{split}
        p_{i}^{\mu} &= (1-x)p_{a}^{\mu} + k_{\perp}^{\mu} - \dfrac{k_{\perp}^{2}}{1-x}\dfrac{n^{\mu}}{2p_{a} n} \, , \\
        p_{a}^{\mu} &= x p_{a}^{\mu} \, ,
        \end{split}
    \end{align}
    with the collinear limit defined as
    \begin{equation}\label{eqn:I_collinear_limit}
        2p_{i}p_{a} = s_{ia} = -\dfrac{k_{\perp}^{2}}{1-x} \, , \quad k_{\perp} \rightarrow 0 \, .
    \end{equation}

    The analogous expression of (\ref{eqn:F_collinear_factorisation})
    for initial-state splitting is
    \begin{equation}\label{eqn:I_collinear_factorisation}
        |\mathcal{M}_{n+1}|^{2} \rightarrow \dfrac{1}{x} \dfrac{1}{p_{i}p_{a}} 4\pi \mu^{2\epsilon} \alpha_{s} \; {}_{n} \langle 1,\ldots,n+1;ai, \ldots | \hat{P}_{ai}(x, k_{\perp}) | 1,\ldots,n+1;ai,\ldots \rangle_{n} \, ,
    \end{equation}
    where the replacement of partons $a$ and $i$ have been made
    explicit by the presence of the parton $ai$ in the colour + helicity vector.
    The specific type of parton $ai$ depends upon the types of $a$ and $i$.

    Similar to (\ref{eqn:soft_factorisation}), (\ref{eqn:F_collinear_factorisation})
     and (\ref{eqn:I_collinear_factorisation}) are only
    true in the strict collinear limit. Away from these
    limits care has to be taken to conserve momenta
    via the use of momenta mappings.

\subsection{Factorisation of colour-ordered amplitudes}
    The factorisation expressed in (\ref{eqn:soft_factorisation}),
    (\ref{eqn:F_collinear_factorisation}), and (\ref{eqn:I_collinear_factorisation})
    were not exact because of colour and spin correlations.

    It can be shown that the factorisation of matrix
    elements in the soft and collinear limits becomes
    exact once the colour structure of the gauge group
    is separated from the kinematics.

    For the scope of this thesis it is sufficient
    to consider colour decomposition of electron-positron
    annihilation. In general any QCD amplitude can be
    colour decomposed. Consider the process $e^{+}e^{-} \rightarrow q\bar{q} + n g$,
    the amplitude can be written as a product of hadronic and
    leptonic currents \cite{Campbell:1998nn} (on which this discussion is based on)
    \begin{equation}\label{eqn:epem_amplitude}
        \mathcal{M}(q_{1}, \bar{q}_{2}; 1, \ldots, n) = \hat{\mathcal{S}}_{\mu}^{n+2}(q_{1}; 1,\ldots,n; \bar{q}_{2})V^{\mu} \, ,
    \end{equation}
    where the leptonic current is $V^{\mu}$ and the hadronic
    current is
    \begin{equation}\label{eqn:hadronic_current}
        \hat{\mathcal{S}}_{\mu}^{n+2}(q_{1};1,\ldots,n;\bar{q}_{2}) = \mathrm{i}eg_{s}^{n} \sum_{P(1,\ldots,n)}(T^{a_{1}} \ldots T^{a_{n}})_{c_{1}c_{2}}S_{\mu}(q_{1};1,\ldots,n;\bar{q}_{2}) \, .
    \end{equation}
    In this expression we have the electromagnetic gauge coupling
    $e$ and the strong gauge coupling $g_{s}$ appearing. The colour
    structure has been factored into a product of fundamental group
    generators where the indices $a_{i} \in {1,\ldots, N_{c}^{2}-1}$
    and $c_{i} \in {1,\ldots,N_{c}}$. This leaves the colour-ordered
    partial amplitude $S_{\mu}(q_{1};1,\ldots,n;\bar{q}_{2})$ depending
    only on kinematical variables, and where the gluons are emitted in an
    ordered fashion from the quarks. The sum over $P(1,\ldots,n)$
    represents a sum over all permutations of gluon emissions
    which accounts for all Feynman diagrams and colour structures.

    Upon squaring the amplitude (\ref{eqn:epem_amplitude}), we get
    \begin{equation}\label{eqn:epem_me2}
        \left|\hat{\mathcal{S}}_{\mu}^{n+2}V^{\mu}\right|^{2} = e^{2}\left(\dfrac{g_{s}^{2}N}{2}\right)^{n}\left(\dfrac{N_{c}^{2}-1}{N_{c}}\right) \sum_{P(1,\ldots,n)} \left(\left|S_{\mu}(q_{1};1,\ldots,n;\bar{q}_{2})V^{\mu}\right|^{2} + \mathcal{O}\left(\dfrac{1}{N_{c}^{2}}\right)\right) \, ,
    \end{equation}
    where the subleading colour terms have been omitted.

    With the amplitude written in terms of the colour-ordered
    partial amplitudes, it is now possible to factorise (\ref{eqn:epem_me2})
    exactly in the soft and collinear limits. In the limit where gluon
    $j$ is soft, we have
    \begin{equation}\label{eqn:CO_soft_factorisation}
        \left|S_{\mu}(q_{1};1,\ldots,i,j,k,\ldots,n;\bar{q}_{2})V^{\mu}\right|^{2} \rightarrow S_{ijk} \left|S_{\mu}(q_{1};1,\ldots,i,k,\ldots,n;\bar{q}_{2})V^{\mu}\right|^{2} \, ,
    \end{equation}
    where the factor
    \begin{equation}\label{eqn:eikonal_factor}
        S_{ijk} = 4\dfrac{s_{ik}}{s_{ij}s_{jk}} \, ,
    \end{equation}
    is the well-known eikonal factor. We see that the colour-ordered
    amplitude on the RHS of (\ref{eqn:eikonal_factor}) has gluon $j$
    removed but the ordering of all hard partons remain unchanged.

    In the collinear limit where partons $i$ and $j$ become collinear
    to form parton $k$, the matrix element factorises as
    \begin{equation}\label{eqn:CO_collinear_factorisation}
        \left|S_{\mu}(q_{1};1,\ldots,i,j,\ldots,n;\bar{q}_{2})V^{\mu}\right|^{2} \rightarrow \dfrac{2}{s_{ij}}P_{ij}(z) \left|S_{\mu}(q_{1};1,\ldots,k,\ldots,n;\bar{q}_{2})V^{\mu}\right|^{2} \, ,
    \end{equation}
    where $P_{ij}(z)$ are the spin-averaged Altarelli-Parisi splitting
    functions (\ref{eqn:AP_kernels}) with the colour factors removed.
    In this expression it is understood that the indices $i$, $j$, and $k$
    have to be self-consistent for the different partonic splittings.
    For partons which are not colour connected (are not neighbouring
    partons in the colour ordered amplitude), there will be no singular
    behaviour as $s_{ij} \rightarrow 0$.

    In (\ref{eqn:CO_soft_factorisation}) and (\ref{eqn:CO_collinear_factorisation})
    the colour-ordered amplitude factorises exactly into a
    colour-ordered amplitude with one parton removed, multiplied
    by a universal singular factor. This singular factor depends on the unresolved
    parton and the two neighbouring hard particles. Interpreting this as
    the two hard particles forming an antenna which radiates the unresolved
    parton gives rise to the antenna functions which will be discussed
    further in Section~\ref{sec:antenna_functions}.

\subsection{One-loop matrix element factorisation}\label{sec:OL_factorisation}
    The discussion of matrix element factorisation so
    far has been focused on tree-level matrix elements.
    It has been shown that one-loop colour-ordered amplitudes
    also factorise in the soft and collinear limits. At the one-loop level,
    there are new universal singular functions and the
    factorisation formula are modified.

    In the soft limit, a one-loop colour-ordered amplitude
    factorises as (cite here)
    \begin{equation}\label{eqn:1L_soft_factorisation}
        M_{n+1}^{(1)}(\ldots, i, j, k, \ldots) \rightarrow S_{ijk}^{(0)} \, M_{n}^{1}(\ldots, i, k, \ldots) + S_{ijk}^{(1)}(\epsilon) \, M_{n}^{0}(\ldots, i, k, \ldots) \, ,
    \end{equation}
    where $S_{ijk}^{(0)}$ is the eikonal factor (\ref{eqn:eikonal_factor}),
    and $S_{ijk}^{(1)}(\epsilon)$ is the one-loop soft radiation function (cite here).

    A similar factorisation formula for the collinear limit is
    \begin{equation}\label{eqn:1L_collinear_factorisation}
        M_{n+1}^{(1)}(\ldots, i, j, \ldots) \rightarrow \dfrac{1}{s_{ij}} \left[ P_{ij}^{(0)}(z) \, M_{n}^{(1)}(\ldots, k, \ldots) + P_{ij}^{(1)}(z, \epsilon) \, M_{n}^{(0)}(\ldots, k, \ldots) \right] \, ,
    \end{equation}
    where $P_{ij}^{(0)}$ are the tree-level splitting functions
    (\ref{eqn:AP_kernels}) and $P_{ij}^{(1)}$
    are the one-loop splitting functions (cite here).

    In both these formulae the structure is of the form:
    tree-level splitting function multipled by a one-loop
    amplitude, plus a one-loop splitting function multiplied
    by a tree-level amplitude as diagrammatically shown
    in Fig ...

    In Section~\ref{sec:antenna_functions} we will give
    a brief overview of antenna functions where these one-loop
    factorisation formulae were used to obtain universal singular functions
    at the one-loop level for squared matrix elements. These
    antenna functions are then applied in the context of NLO
    k-factor emulation in Chapter~\ref{chapter:fame2}.

\section{Subtraction}\label{sec:subtraction}
\subsection{Catani-Seymour dipoles}\label{sec:CS_dipoles}
\subsection{Antenna}\label{sec:antenna_functions}
\end{document}
